# `Course Project 2` Exploratory Data Analysis
================
Anderson H. Uyekita

-   Specialization: [Data Science: Foundations using R
    Specialization](https://www.coursera.org/specializations/data-science-foundations-r)
-   Course: [Exploratory Data
    Analysis](https://www.coursera.org/learn/exploratory-data-analysis)
    -   Instructor: Roger D Peng
-   Week 1:
    -   Start: 2022/05/24
    -   Finish: 2022/05/25
-   [Project Instructions](./instructions.md)
-   [Project README](./README.md)

------------------------------------------------------------------------









# Codebook

## 1. Requirements

It is necessary to install the following packages to reproduce this Course Project 2.

1. `ggplot2`
2. `tidyverse`
3. `magrittr`

It is also mandatory to have access to the internet if you are willing to reproduce it.

### Description of functionality of `plot1.R`

1. Create the folder `data` to keep all files downloaded in it;
2. Download the zipped file using `download.file()`;
3. If the raw data is already downloaded the script will not download any file;
4. Unzipping the txt file (Electric power consumption) into the `data` folder;
5. Loading the txt file using `read.table()` and naming it as raw_database;
6. Update the `Time` columns as a result of merging `Date` and `Time` columns using `strptime()`. Now the `Time` column became POSIXlt objects;
7. Converting `Date` column into Date objects using `as.Date`;
8. Subsetting the raw data between 1/2/2007 and 2/2/2007 to create the Tidy dataset (df_tidy);
9. Plot the histogram of `Global_active_power`;
10. Saving the plot as `png` file;
11. Excluding all variable and turn back the original work directory.

### Description of functionality of `plot2.R`

1. Create the folder `data` to keep all files downloaded in it;
2. Download the zipped file using `download.file()`;
3. If the raw data is already downloaded the script will not download any file;
4. Unzipping the txt file (Electric power consumption) into the `data` folder;
5. Loading the txt file using `read.table()` and naming it as raw_database;
6. Update the `Time` columns as a result of merging `Date` and `Time` columns using `strptime()`. Now the `Time` column became POSIXlt objects;
7. Converting `Date` column into Date objects using `as.Date`;
8. Subsetting the raw data between 1/2/2007 and 2/2/2007 to create the Tidy dataset (df_tidy);
9. Plot the line graphic of `Global_active_power` by `1/2/2007 and 2/2/2007`;
10. Saving the plot as `png` file;
11. Excluding all variable and turn back the original work directory.

### Description of functionality of `plot3.R`

1. Create the folder `data` to keep all files downloaded in it;
2. Download the zipped file using `download.file()`;
3. If the raw data is already downloaded the script will not download any file;
4. Unzipping the txt file (Electric power consumption) into the `data` folder;
5. Loading the txt file using `read.table()` and naming it as raw_database;
6. Update the `Time` columns as a result of merging `Date` and `Time` columns using `strptime()`. Now the `Time` column became POSIXlt objects;
7. Converting `Date` column into Date objects using `as.Date`;
8. Subsetting the raw data between 1/2/2007 and 2/2/2007 to create the Tidy dataset (df_tidy);
9. Plot the line graphic of `Sub_metering_1` by `1/2/2007 and 2/2/2007`;
10. Add to the graphic above the `Sub_metering_2` and `Sub_metering_3`;
11. Insert a legend;
12. Saving the plot as `png` file;
13. Excluding all variable and turn back the original work directory.

### Description of functionality of `plot4.R`

1. Create the folder `data` to keep all files downloaded in it;
2. Download the zipped file using `download.file()`;
3. If the raw data is already downloaded the script will not download any file;
4. Unzipping the txt file (Electric power consumption) into the `data` folder;
5. Loading the txt file using `read.table()` and naming it as raw_database;
6. Update the `Time` columns as a result of merging `Date` and `Time` columns using `strptime()`. Now the `Time` column became POSIXlt objects;
7. Converting `Date` column into Date objects using `as.Date`;
8. Subsetting the raw data between 1/2/2007 and 2/2/2007 to create the Tidy dataset (df_tidy);
9. Set the global graphic configuration as two-by-two (`par(mfrow = c(2, 2))`);
10. Plot the line graphic of `Global_active_power` by `1/2/2007 and 2/2/2007` at the top left;
11. Plot the line graphic of `Voltage` by `1/2/2007 and 2/2/2007` at the top right;
12. Plot the line graphic of `Sub_metering_1` by `1/2/2007 and 2/2/2007` and  `Sub_metering_2` and `Sub_metering_3` at bottom left;
13. Plot the line graphic of `Global_reactive_power` by `1/2/2007 and 2/2/2007` at the bottom right;
14. Saving the plot as `png` file;
15. Excluding all variable and turn back the original work directory.

## 2. Output and Input Details

Here will be described some aspects of output and inputs.

### 2.1. Inputs

```{r inputs,cache=TRUE,message=FALSE,warning=FALSE,include=FALSE}
# 1. Create a data directory
if(!base::file.exists("data")) {
    base::dir.create("data")
    }

# 2. Download files and store it in data directory.
if(!base::file.exists("./data/FNEI_data.zip")){
    utils::download.file(url = "https://d396qusza40orc.cloudfront.net/exdata%2Fdata%2FNEI_data.zip",
                         destfile = "./data/FNEI_data.zip")
    }

# 2.1. Unzipping the FNEI_data.zip file.
if(!base::file.exists("./data/unzipped/Source_Classification_Code.rds") | !base::file.exists("./data/unzipped/summarySCC_PM25.rds")){
    utils::unzip(zipfile = "./data/FNEI_data.zip",
                 exdir = "./data/unzipped/",
                 list = FALSE,
                 overwrite = TRUE)
    }
```

Course Project 2 will only use one zipped file, the `FNEI_data.zip`. It has 2 (two) `.rds` compressed files:

* `summarySCC_PM25.rds`, and;
* `Source_Classification_Code.rds`.

#### 2.1.1. NEI

```{r reading-rds-nei,cache=TRUE,warning=FALSE,message=FALSE,include=FALSE}
NEI <- readRDS("./data/unzipped/summarySCC_PM25.rds")
```

The data from `summarySCC_PM25.rds` is already tidy, and following the assignment instruction, I have named this object `NEI`.

* Size: `r base::format(pryr::object_size(NEI), big.mark = ",")` Bytes (`r base::format(pryr::object_size(NEI)/1048576, big.mark = ",", nsmall  = 1)` Megabytes)
* Rows: `r base::format(base::dim(NEI)[1], big.mark = ",")`, and;
* Columns: `r base::format(base::dim(NEI)[2], big.mark = ",")`.

I have calculated the object size in Bytes using the `object_size()` from the `pryr` package.

The columns variables in the NEI dataset: `r base::colnames(NEI)`

**Head**

```{r nei-head,cache=TRUE,warning=FALSE,message=FALSE,include=TRUE,echo=FALSE}
library(kableExtra)

utils::head(NEI) %>% kableExtra::kbl() %>% kableExtra::kable_styling() 
```

**Strucuture**

```{r nei-str,cache=TRUE,warning=FALSE,message=FALSE,include=TRUE,echo=FALSE}
utils::str(NEI)
```

#### 2.1.2. SCC

```{r reading-rds-scc,cache=TRUE,warning=FALSE,message=FALSE,include=FALSE}
SCC <- readRDS("./data/unzipped/Source_Classification_Code.rds")
```

The data from `Source_Classification_Code.rds` is already tidy. Following the assignment instruction, I have named this object `SCC`.

* Size: `r base::format(pryr::object_size(SCC), big.mark = ",")` Bytes (`r base::format(pryr::object_size(SCC)/1048576, big.mark = ",", nsmall  = 1)` Megabytes)
* Rows: `r base::format(base::dim(SCC)[1], big.mark = ",")`, and;
* Columns: `r base::format(base::dim(SCC)[2], big.mark = ",")`.

I have calculated the object size in Bytes using the `object_size()` from the `pryr` package.

The columns variables in the SCC dataset: `r colnames(SCC)`

**Head**

```{r scc-head,cache=TRUE,warning=FALSE,message=FALSE,include=TRUE,echo=FALSE}
library(kableExtra)

utils::head(SCC) %>% kableExtra::kbl() %>% kableExtra::kable_styling() 
```

**Strucuture**

```{r scc-str,cache=TRUE,warning=FALSE,message=FALSE,include=TRUE,echo=FALSE}
utils::str(SCC)
```

### 2.2. Output

The results of this assignment are 6 (six) files `.R` and 6 (six) files `png`.

**EI.Sector**

There are 59 types of `EI.Sector`:

```{r,cache=TRUE,warning=FALSE,message=FALSE,include=TRUE,echo=FALSE}
unique(SCC$EI.Sector)
```

The EI.Sector related to Coal:

```{r,cache=TRUE,warning=FALSE,message=FALSE,include=TRUE,echo=FALSE}
unique(SCC$EI.Sector)[grepl(unique(SCC$EI.Sector), pattern = "Coal|coal")]
```