---
title: '`Course Project 2` Exploratory Data Analysis'
author: '`r if(knitr::is_html_output()) {"&#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Anderson H Uyekita"} else {NULL}`'
output:
  github_document: default
params:
  author: "Anderson H Uyekita"
  course: "Exploratory Data Analysis"
  course_url: "https://www.coursera.org/learn/exploratory-data-analysis"
  specialization: "Data Science: Foundations using R Specialization"
  specialization_url: "https://www.coursera.org/specializations/data-science-foundations-r"
  instructor: "Roger D Peng"
  week: "Week 4"
  course_start: !r lubridate::ymd("2022/06/15")
  course_finish: !r lubridate::ymd("2022/06/19")
  slug: 'Week%204'
  rpubs: 'https://rpubs.com/AndersonUyekita/course-project-2_exploratory-data-analysis'
  project_repo: 'https://github.com/AndersonUyekita/exploratory-data-analysis_course-project-2'
  codebook: 'https://github.com/AndersonUyekita/exploratory-data-analysis_course-project-2/blob/master/codebook.md'
  instructions: 'https://github.com/AndersonUyekita/exploratory-data-analysis_course-project-2/blob/master/instructions.md'
always_allow_html: true
---

`r if(!knitr::is_html_output()) {sprintf(fmt = "* &#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Author: %s", params$author)}`
`r sprintf(fmt = "* &#x1f4da; Specialization: [%s](%s){target='_blank' rel='noopener'}", params$specialization, params$specialization_url)`
`r sprintf(fmt = "* &#x1f4d6; Course: [%s](%s){target='_blank' rel='noopener'}", params$course, params$course_url)`
    `r sprintf(fmt = "* &#x1F9D1;&#x200d;&#x1F3EB; Instructor: %s", params$instructor)`
`r sprintf(fmt = "* &#x1F4C6; %s", params$week)`
    `r sprintf(fmt = "* &#x1F6A6; Start: %s", format(params$course_start, "%A, %d %B %Y"))`
    `r sprintf(fmt = "* &#x1F3C1; Finish: %s", format(params$course_finish, "%A, %d %B %Y"))`
`r if(!knitr::is_html_output()) {sprintf(fmt = "* &#x1F30E; Rpubs: [Interactive Document](%s)", params$rpubs)}else{sprintf(fmt = "* &#x1F4E6; Github Repository: [Static Document](%s){target='_blank' rel='noopener'}", params$project_repo)}`
`r if(!knitr::is_html_output()) {sprintf(fmt = "* &#x1F4CB; Instructions: [Project Instructions](%s)", "./instructions.md")}else{sprintf(fmt = "* &#x1F4CB; Instructions: [Project Instructions](%s){target='_blank' rel='noopener'}", params$instructions)}`
`r if(!knitr::is_html_output()) {sprintf(fmt = "* &#x1F4C4; README: [README.md](%s)", "./README.md")}else{sprintf(fmt = "* &#x1F4C4; README: [README.md](%s){target='_blank' rel='noopener'}", params$project_repo)}`

## Codebook

1. [Requirements](#requirements)
2. [Scripts](#scripts)
    1. [Plot 1](#plot1)
    1. [Plot 2](#plot2)
    1. [Plot 3](#plot3)
    1. [Plot 4](#plot4)
    1. [Plot 5](#plot5)
    1. [Plot 6](#plot6)
3. Output and Input Details
    1. [Inputs](#inputs)
        1. [NEI](#nei)
        2. [SCC](#scc)
    2. [Output](#output)

### 1. Requirements [](){id="requirements"}

It is necessary to install the following packages to reproduce this Course Project 2.

1. `ggplot2`
2. `tidyverse`
3. `magrittr`
4. `cowplot`

Also, those following scripts will use the standard packages from R.

1. `base`
2. `graphic`
3. `utils`
4. `grDevices`

It is mandatory to have access to the internet if you are willing to reproduce it.

### 2. Scripts [](){id="scripts"}

This Course Project 2 required the development of 6 scripts to plot and export six png files. 

#### 2.1. Description of functionality of `plot1.R` [](){id="plot1"}

1. Create the folder `data` to keep all files downloaded in it;
2. Download the zipped file (FNEI_data.zip) using `download.file()`;
3. If the raw data is already downloaded, the script will not download any file;
4. Unzipping the `FNEI_data.zip` into the `data` folder;
5. Loading the `summarySCC_PM25.rds` file using `base::readRDS` and naming it as `NEI`;
6. Loading the `Source_Classification_Code.rds` file using `base::readRDS` and naming it as `SCC`;
7. Summarizing the NEI dataset grouping by year and then calculating the Total emissions of PM2.5;
8. Plotting a barplot using the base graphic system;
9. Saving the plot as a `png` file;

#### 2.2. Description of functionality of `plot2.R` [](){id="plot2"}

1. Creating the folder `data` to keep all files downloaded in it;
2. Downloading the zipped file (FNEI_data.zip) using `download.file()`;
3. If the raw data is already downloaded, the script will not download any file;
4. Unzipping the `FNEI_data.zip` into the `data` folder;
5. Loading the `summarySCC_PM25.rds` file using `base::readRDS` and naming it as `NEI`;
6. Loading the `Source_Classification_Code.rds` file using `base::readRDS` and naming it as `SCC`;
7. Subsetting the NEI dataset to filter only fips equal to 24510 (Baltimore City);
8. Summarizing the previous dataset grouping by year and then calculating the Total emissions of PM2.5;
9. Plotting a line plot using the base graphic system;
10. Saving the plot as a `png` file;

#### 2.3. Description of functionality of `plot3.R` [](){id="plot3"}

1. Creating the folder `data` to keep all files downloaded in it;
2. Downloading the zipped file (FNEI_data.zip) using `download.file()`;
3. If the raw data is already downloaded, the script will not download any file;
4. Unzipping the `FNEI_data.zip` into the `data` folder;
5. Loading the `summarySCC_PM25.rds` file using `base::readRDS` and naming it as `NEI`;
6. Loading the `Source_Classification_Code.rds` file using `base::readRDS` and naming it as `SCC`;
7. Subsetting the NEI dataset to filter only fips equal to 24510 (Baltimore City);
8. Summarizing the previous dataset grouping by year and then calculating the Total emissions of PM2.5;
9. Plotting a line plot using the `ggplot2` package;
10. Setting the aesthetic from geom_line as `color = type`;
10. Saving the plot as a `png` file;

#### 2.4. Description of functionality of `plot4.R` [](){id="plot4"}

1. Creating the folder `data` to keep all files downloaded in it;
2. Downloading the zipped file (FNEI_data.zip) using `download.file()`;
3. If the raw data is already downloaded, the script will not download any file;
4. Unzipping the `FNEI_data.zip` into the `data` folder;
5. Loading the `summarySCC_PM25.rds` file using `base::readRDS` and naming it as `NEI`;
6. Loading the `Source_Classification_Code.rds` file using `base::readRDS` and naming it as `SCC`;
7. Subsetting the SCC dataset to filter the Coal source of PM2.5 found in EI.Sector variable;
8. Subsetting the NEI dataset to filter only the given SCC from the previous dataset;
9. Merging both previous datasets to create a new column about EI.Sector;
8. Summarizing the previous dataset grouping by year and EI.The sector then calculating the Total emissions of PM2.5;
9. Plotting a stacked bar plot using the `ggplot2` package;
10. Setting the geom_bar to have `position = "stack"`;
10. Saving the plot as a `png` file;

#### 2.5. Description of functionality of `plot5.R` [](){id="plot5"}

1. Creating the folder `data` to keep all files downloaded in it;
2. Downloading the zipped file (FNEI_data.zip) using `download.file()`;
3. If the raw data is already downloaded, the script will not download any file;
4. Unzipping the `FNEI_data.zip` into the `data` folder;
5. Loading the `summarySCC_PM25.rds` file using `base::readRDS` and naming it as `NEI`;
6. Loading the `Source_Classification_Code.rds` file using `base::readRDS` and naming it as `SCC`;
7. Subsetting the NEI dataset to filter only fips equal to 24510 (Baltimore City) and type equal to ON-ROAD;
8. Summarizing the previous dataset grouping by year and then calculating the Total emissions of PM2.5;
9. Plotting a bar plot using the `ggplot2` package;
10. Saving the plot as a `png` file;

#### 2.6. Description of functionality of `plot6.R` [](){id="plot6"}

1. Creating the folder `data` to keep all files downloaded in it;
2. Downloading the zipped file (FNEI_data.zip) using `download.file()`;
3. If the raw data is already downloaded, the script will not download any file;
4. Unzipping the `FNEI_data.zip` into the `data` folder;
5. Loading the `summarySCC_PM25.rds` file using `base::readRDS` and naming it as `NEI`;
6. Loading the `Source_Classification_Code.rds` file using `base::readRDS` and naming it as `SCC`;
7. Subsetting the NEI dataset to filter only fips equal to 24510 or 06037 (Baltimore City and Los Angeles County) and type equal to ON-ROAD;
8. Creating an auxiliary dataset to store the fips and city;
9. Merging both previous datasets to create a new column about the city;
10. Summarizing the previous dataset grouping by year and then calculating the Total emissions of PM2.5;
11. Plotting a bar plot using the `ggplot2` package;
12. Saving the plot as a `png` file;

## 3. Output and Input Details

Here will be described some aspects of output and inputs.

### 3.1. Inputs [](){id="inputs"}

```{r inputs,cache=TRUE,message=FALSE,warning=FALSE,include=FALSE}
# 1. Create a data directory
if(!base::file.exists("data")) {
    base::dir.create("data")
    }

# 2. Download files and store it in data directory.
if(!base::file.exists("./data/FNEI_data.zip")){
    utils::download.file(url = "https://d396qusza40orc.cloudfront.net/exdata%2Fdata%2FNEI_data.zip",
                         destfile = "./data/FNEI_data.zip")
    }

# 2.1. Unzipping the FNEI_data.zip file.
if(!base::file.exists("./data/unzipped/Source_Classification_Code.rds") | !base::file.exists("./data/unzipped/summarySCC_PM25.rds")){
    utils::unzip(zipfile = "./data/FNEI_data.zip",
                 exdir = "./data/unzipped/",
                 list = FALSE,
                 overwrite = TRUE)
    }
```

Course Project 2 will only use one zipped file, the `FNEI_data.zip`. It has 2 (two) `.rds` compressed files:

* `summarySCC_PM25.rds`, and;
* `Source_Classification_Code.rds`.

#### 3.1.1. NEI [](){id="nei"}

```{r reading-rds-nei,cache=TRUE,warning=FALSE,message=FALSE,include=FALSE}
NEI <- readRDS("./data/unzipped/summarySCC_PM25.rds")
```

The data from `summarySCC_PM25.rds` is already tidy, and following the assignment instruction, I have named this object `NEI`.

* Size: `r base::format(pryr::object_size(NEI), big.mark = ",")` Bytes (`r base::format(pryr::object_size(NEI)/1048576, big.mark = ",", nsmall  = 1)` Megabytes)
* Rows: `r base::format(base::dim(NEI)[1], big.mark = ",")`, and;
* Columns: `r base::format(base::dim(NEI)[2], big.mark = ",")`.

I have calculated the object size in Bytes using the `object_size()` from the `pryr` package.

The columns variables in the NEI dataset: `r base::colnames(NEI)`

**Head**

```{r nei-head,cache=TRUE,warning=FALSE,message=FALSE,include=TRUE,echo=FALSE}
library(kableExtra)

utils::head(NEI) %>% kableExtra::kbl() %>% kableExtra::kable_styling() 
```

**Strucuture**

```{r nei-str,cache=TRUE,warning=FALSE,message=FALSE,include=TRUE,echo=FALSE}
utils::str(NEI)
```

#### 3.1.2. SCC [](){id="scc"}

```{r reading-rds-scc,cache=TRUE,warning=FALSE,message=FALSE,include=FALSE}
SCC <- readRDS("./data/unzipped/Source_Classification_Code.rds")
```

The data from `Source_Classification_Code.rds` is already tidy. Following the assignment instruction, I have named this object `SCC`.

* Size: `r base::format(pryr::object_size(SCC), big.mark = ",")` Bytes (`r base::format(pryr::object_size(SCC)/1048576, big.mark = ",", nsmall  = 1)` Megabytes)
* Rows: `r base::format(base::dim(SCC)[1], big.mark = ",")`, and;
* Columns: `r base::format(base::dim(SCC)[2], big.mark = ",")`.

I have calculated the object size in Bytes using the `object_size()` from the `pryr` package.

The columns variables in the SCC dataset: `r colnames(SCC)`

**Head**

```{r scc-head,cache=TRUE,warning=FALSE,message=FALSE,include=TRUE,echo=FALSE}
library(kableExtra)

utils::head(SCC) %>% kableExtra::kbl() %>% kableExtra::kable_styling() 
```

**Strucuture**

```{r scc-str,cache=TRUE,warning=FALSE,message=FALSE,include=TRUE,echo=FALSE}
utils::str(SCC)
```

### 3.2. Output [](){id="output"}

The results of this assignment are 6 (six) files `.R` and 6 (six) files `png`.

**EI.Sector**

There are 59 types of `EI.Sector`:

```{r,cache=TRUE,warning=FALSE,message=FALSE,include=TRUE,echo=FALSE}
unique(SCC$EI.Sector)
```

The EI.Sector related to Coal:

```{r,cache=TRUE,warning=FALSE,message=FALSE,include=TRUE,echo=FALSE}
unique(SCC$EI.Sector)[grepl(unique(SCC$EI.Sector), pattern = "Coal|coal")]
```