---
title: '`Course Prokect 2` Exploratory Data Analysis'
author: '`r if(knitr::is_html_output()) {"&#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Anderson H Uyekita"} else {NULL}`'
output:
  github_document: default
params:
  author: "Anderson H Uyekita"
  course: "Exploratory Data Analysis"
  course_url: "https://www.coursera.org/learn/exploratory-data-analysis"
  specialization: "Data Science: Foundations using R Specialization"
  specialization_url: "https://www.coursera.org/specializations/data-science-foundations-r"
  instructor: "Roger D Peng"
  week: "Week 4"
  course_start: !r lubridate::ymd("2022/06/15")
  course_finish: !r lubridate::today()
  slug: 'Week%204'
---

```{r setup, include=FALSE,message=FALSE,warning=FALSE,cache=TRUE}
# Defining document chunks options
knitr::opts_chunk$set(echo = TRUE,cache=TRUE,message=FALSE,warning=FALSE)

# Loading required libraries
library(rmarkdown)
library(lubridate)
library(ggplot2)
```

* `r if(!knitr::is_html_output()) {sprintf(fmt = "&#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Author: %s", params$author)}`
* `r sprintf(fmt = "&#x1f4da; Specialization: [%s](%s){target='_blank' rel='noopener'}", params$specialization, params$specialization_url)`
* `r sprintf(fmt = "&#x1f4d6; Course: [%s](%s){target='_blank' rel='noopener'}", params$course, params$course_url)`
    * `r sprintf(fmt = "&#x1F9D1;&#x200d;&#x1F3EB; Instructor: %s", params$instructor)`
* `r sprintf(fmt = "&#x1F4C6; %s", params$week)`
    * `r sprintf(fmt = "&#x1F6A6; Start: %s", format(params$course_start, "%A, %d %B %Y"))`
    * `r sprintf(fmt = "&#x1F3C1; Finish: %s", format(params$course_finish, "%A, %d %B %Y"))`
* `r if(!knitr::is_html_output()) {sprintf("&#x1F30E; Rpubs: [Interactive Document](%s)", params$rpubs)}else{sprintf("&#x1F4E6; Github Repository: [Static Document](%s){target='_blank' rel='noopener'}", params$project_repo)}`

--------------------------------------------------------------------------------

## Sinopsis



--------------------------------------------------------------------------------

## 1. Project Course 2 Assignment

Project Course 2 consists of answering six questions. Each question assignment asks to create a single plot to answer it.

#### Plot 1

> Question 1: Have total emissions from PM2.5 decreased in the United States from 1999 to 2008?

Using the **base** plotting system, make a plot showing the *total* PM2.5 emission from all sources for each of the years 1999, 2002, 2005, and 2008.

```{r plot-1,cache=TRUE,message=FALSE,warning=FALSE,echo=FALSE}
source(file = "plot1.R")
```

[](./plot1.png)

Over the years, the fine particulate matter, the so-called PM2.5, has decreased from 7.3 million tons to 3.5 million tons, representing a 52.7% reduction.

#### Plot 2

> Question 2: Have total emissions from PM2.5 decreased in the **Baltimore City**, Maryland (`fips == "24510"`) from 1999 to 2008?

Use the **base** plotting system to make a plot answering this question.

```{r plot-2,cache=TRUE,message=FALSE,warning=FALSE,echo=FALSE}
source(file = "plot2.R")
```

[](./plot2.png)

Based on Plot 2, the fine particulate matter (PM2.5) decreased by 43.1% from 1999 to 2008. However, in 2005 the PM2.5 of Baltimore City surpassed the 2002 index, showing a disturbance in the trend of PM 2.5 decreasing.

#### Plot 3

> Question 3: Of the four types of sources indicated by the `type` (point, nonpoint, onroad, nonroad) variable, which of these four sources have seen decreases in emissions from 1999–2008 for **Baltimore City**? Which have seen increases in emissions from 1999–2008?



Use the **ggplot2** plotting system to make a plot answer this question.




#### Plot 4

Across the United States, how have emissions from coal combustion-related sources changed from 1999–2008?




#### Plot 5

How have emissions from motor vehicle sources changed from 1999–2008 in **Baltimore City**?



#### Plot 6

Compare emissions from motor vehicle sources in Baltimore City with emissions from motor vehicle sources in **Los Angeles County**, California (`fips == "06037"`). Which city has seen greater changes over time in motor vehicle emissions?


## 2. 
